---
title: "p8105_hw5_mgw2144"
output: github_document
date: "2023-11-08"
---

```{r message=FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = ("viridis"))

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1 


# Problem 2
## Reading the data and creating a dataframe with the map function

```{r message=FALSE, warning=FALSE}
study_data = 
  tibble(
    files = list.files('data/study_data'),
    path = str_c("data/study_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest() 
```

## Tidying the data

```{r}

study_data_tidy = 
  study_data %>% 
  separate(files, into = c("control_arm", "subject_ID", "csv")) %>% 
  mutate(
    control_arm = ifelse(control_arm == "con", "control", control_arm),
    control_arm = ifelse(control_arm == "exp", "experimental", control_arm)
  ) %>% 
  pivot_longer(
    cols = starts_with("week"),
    names_prefix = "week_",
    names_to = "week_number",
    values_to = "observation"
  ) %>% 
  mutate(
     week_number = as.numeric(week_number)) %>% 
  select(control_arm, subject_ID, week_number, observation)

```

## Plotting the tidy data 

```{r}

study_plot = 
  study_data_tidy %>% 
  ggplot(aes(x = week_number, y = observation, colour = subject_ID)) +
  geom_point() +
  geom_line() +
  facet_grid(.~control_arm) +
  labs(
    x = "Week",
    y = "Observation", 
    title = "Observations by week for each subject in both control arms" 
  )

study_plot

```

The control arm appears to have observations steadily between the range of approximately -2.3 and 4.3 for all 8 weeks and for all 10 subjects with the lowest observation value seen in week 6 for subject 5 and the highest observation value seen in week 5 for subject 10. The experimental arm shows a steady positive/increasing trend for the observations values in all subjects across the 8 week period. The lowest observation value for the experimental arm was in subject 1 in week 1, and the highest observational value was in week 9 for subject 5. This indicates a positive relationship between observation value and week (time) in the experimental arm that we do not see in the control arm. 

# Problem 3 

## Creating function and map
```{r}
sim_function = function(n = 30, mu_val = c(0, 1, 2, 3, 4, 5, 6), sigma = 5, alpha = 0.05, num_datasets = 5000) {
  
  results = map_df(mu_val, function(mu) {
    
    datasets = replicate(num_datasets, rnorm(n, mean = mu, sd = sigma), simplify = FALSE)
    
    map_df(datasets, function(dataset) {
      t_test_result = t.test(dataset, mu = mu)
      tidy_results = broom::tidy(t_test_result)
      data.frame(mu = mu, mu_est = tidy_results$estimate, p_value = tidy_results$p.value)
    })
  })
  
  return(results)
}

result_df = sim_function()

```

## Plot 1: Power vs. True Mean 

```{r}
plot_1_df =
  result_df %>%
  group_by(mu) %>%
  summarise(power = mean(p_value < 0.05))

plot_1=
ggplot(plot_1_df, aes(x = mu, y = power)) +
  geom_line() +
  geom_point() +
  labs(
    x = "True mu value",
    y = "Power of the Test",
    title = "Association between effect size and power"
  )

plot_1
```


There appears to be a non-linear association between effect size and power based on the graph. Each time that we run the simulated data the graph appears different, with a non-linear association each time. 


## Plot 2: Average estimate of mu hat compared to true mu value 

```{r}
plot_2_df =
result_df %>%
  group_by(mu) %>%
  summarise(avg_mu_hat = mean(mu_est))
  
plot_2 = 
  plot_2_df %>% 
  ggplot(aes(x = mu, y = avg_mu_hat)) +
  geom_point()+
  geom_line()+
  labs(
    title = "Average estiamte of mu hat vs. the true mu value",
    x = "True mu value",
    y = "Estimated mu value"
  )

plot_2

```

There is a very strong positive linear relationship between the true mu values and the average estimated mu values created from the simulation. This makes sense as when we run a higher number of samples, our average estimated means will mimic the true mean value. 

## Plot 3: Avergate estimate of mu in samples where the null is rejected compared to the true mean value

```{r}
plot_3_df = 
  result_df %>%
  filter(p_value < 0.05) %>%
  group_by(mu) %>%
  summarise(avg_mu_hat_rej = mean(mu_est))

plot_3 =
  ggplot() +
  geom_line(data = plot_2_df, aes(x = mu, y = avg_mu_hat), color = "green")+
  geom_point(data = plot_2_df, aes(x = mu, y = avg_mu_hat), color = "green")+
  geom_line(data = plot_3_df, aes(x = mu, y = avg_mu_hat_rej), color = "blue") +
  geom_point(data = plot_3_df, aes(x = mu, y = avg_mu_hat_rej), color = "blue") +
  labs(
    x = "True value of mu",
    y = "Average estimate of mu",
    title = "Average estimate of mu for samples where null hypothesis was rejected"
  )

plot_3
```
 
The sample average of mu hat across tests for which the null is rejected is approximately equal to the true value of mu. As we run more samples, the average values of the estimated mus more closely approximates the true mu values. The differences seen in the graph is due to random sampling variability in the results.
